Hydras, Ordinals \& Co. \\ A library in Coq of entertaining formal mathematics

### 2021-10-12 doc/JFLA2022/paper.tex
pass over intro

Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}. However, after a key mathematical definition or result is added to a library, it needs be well \emph{documented} and \emph{maintained}.

Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions. Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms to beginners~\cite{MCB}. The former tends to be terse and dense, while the latter is more long-winded and exhaustive.

As a library changes and expands, its documentation needs to be made consistent and complete. Authors use a wide range of techniques, including literate programming and custom tools, to check code snippets and generate output from the proof assistant.

>>>>

Libraries of formalized mathematics based on proof assistants, such as \coq, Lean, and Isabelle/HOL, are continually growing in size and scope.
For example, the core projects in the Mathematical Components family for Coq amount to more than 160,000 lines of code and 11,000 lemmas~\cite{Nie2020a}, while Lean's Mathlib consists of more than 140,000 lines of code and 34,000 declarations~\cite{Mathlib}. However, after a key mathematical definition or result is added to a library, it needs be \emph{documented} and \emph{maintained}.

Maintenance includes not only adaptation to changes in new proof assistant versions, but also reorganization to ease future maintenance effort and accommodate new contributions. Documentation is usually two-fold: \emph{source code comments} that describe specific definitions and results in-line and \emph{books} that carefully introduce the library and its idioms~\cite{MCB}. The former tends to be terse and dense and serves experienced users, while the latter is more long-winded and exhaustive and serves beginners. As a library changes and expands, all its documentation needs to be made consistent and complete. Authors use a wide range of techniques, including literate programming~\cite{Knuth1984} and custom tools, to pretty-print and check source code snippets and generate proof assistant output.

While books that document proof assistant libraries are valuable to beginners, the purpose of most current books is mainly instrumental, i.e., to teach a certain topic or technique. We believe that such books can become \emph{ends-in-themselves}---not just sources of learning, but also of inspiration and entertainment.
